---
title: "Data import"
output: github_document
Date: 9/20/2022
---


```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

##Read in some data

Read in the litters dataset.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

##Take a look at the data

Printing in the console.

```{r}
litters_df 
```

```{r}
tail(litters_df) 
```

```{r}
skimr::skim(litters_df)
```

##Options to read csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

##Other file formats

Read in an excel file.

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

Read in a sas file

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

##Comparison with Base R



# Data maniputation

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

Choose some columns and not others.

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

Renaming columns

```{r}
select(litters_df, GROUP = group)
```

```{r}
rename(litters_df, GROUP = group)
```

Select helpers

```{r}
select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, litter_number, everything())
```

```{r}
relocate(litters_df, litter_number)
```

## 'filter'

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

Whether =20?
```{r}
filter(litters_df, gd_of_birth == 20)
```

```{r}
filter(litters_df, !(gd_of_birth == 20))
filter(litters_df, gd_of_birth != 20)
```

```{r}
filter(litters_df, group == "Mod8")
```

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## 'mutate'

str_to_lower: lowercase

```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight -gd0_weight,
  group = str_to_lower(group))
```

##'arrange'

```{r}
arrange(litters_df, pups_born_alive)
```

## '%>%'

```{r}
litters_data_row = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_row)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_without_missing = drop_na(litters_data_selected, gd0_weight)
```

use the pipe operator instead

```{r}
litters_df =
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  drop_na(gd0_weight)
```



#Tidy data

##pivot_longer

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

wide format to long format

```{r}
pulse_data_tidy =
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
```

rewrite, combine, and extend (to add a mutate)

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>%
  mutate(visit = recode(visit, "bl" = "00m"))
```

## 'pivot_wider'

make up some data

```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result %>% 
  pivot_wider(
    names_from = "time",
    values_from = "mean"
  )
```

## binding rows

Using the lotR data

First step: import data

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")
```

Bind all the rows together

```{r}
lotr_tidy =
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  relocate(movie) %>% 
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "words"
  )
```

